---
title: "4.Syndrome_scales_rsfmri"
author: "Lorenza Dall'Aglio"
date: "2/22/2022"
output: html_document
---

# Cross-lagged panel models rsfMRI - psychiatric problems

*Project* rsfmri - psychiatric problems
*Author* Lorenza Dall'Aglio (l.dallaglio@erasmusmc.nl; lorenza.dallaglio1@gmail.com)
*Dataset* the Generation R Study
*Date* 22 Feb 2022

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

## Data prep

```{r}

rm(list=ls())

# load env 

source("/PUT/RELEVANT/PATH/0.Source_file.R")

# load data 

dd <- readRDS(paste0(indata, "genr_data_main_rsfMRI_psych.rds"))

summary(dd)

tab <- "/PUT/RELEVANT/PATH/"

```

```{r}

####
# factor scaling 
####
# this is needed so that the variables are more or less on the same order of magnitude
# FA/MD values are generally 100 times smaller than behavioural ones. --> use scale factor of 100 

scale_factor = 100


mdts <- grep("mdt", names(dd), value = T)

dd[mdts] <- dd[mdts] / scale_factor



####
# Recoding of vars for lavaan
####
# Before running lavaan, variables need to be coded in a certain way if they are not continuous
# this varies depending on whether the var is exogeneous (ind) or endogeneous (dep)
# more info at https://lavaan.ugent.be/tutorial/cat.html


### Make ordered variables, continuous/numeric ###
# maternal education only here

dd$mat_edu_num <- as.numeric(dd$mat_edu)



### dummy coding for binary/nominal variables ### 
# to make dummy cols from factors or character vars you can use dummy_cols()
# NB check which vars have factors or chr 
# id is chr - change it just temporarily so it is not transformed into a dummy 
# only nominal vars here are ethnicity and sex

rownames(dd) <- dd$id
dd$id <- NULL

dd2 <- fastDummies::dummy_cols(dd)   
# NB the function dichotomises everything w/o setting ref cateogory so you need to take care of that


# delete reference vars 
# and dummies with NAs 

dd2$ethn_dutch <- NULL
dd2$ethn_NA <- NULL
dd2$sex_girl <- NULL
```


## Model specification 
We need to test the relation of each MDT and of NT with each syndrome scale

Because we have 5 MDT states + a var for NT, we have a total of 6 rsfMRI measures to be tested with the 8 syndrome scales

For a total of 48 models and 96 lagged paths of interest 


### general model specification 

Because of the many models, I created general model specifications (one for int, one for ext)  so that
the model specifications can then be adapted to each syndrome
The object `phen` is where the syndrome scale measures will go 
The object `state` is where the resting state measures will go 

```{r}

# general model 

m_general  <- '

# lagged and stability 
phen_t2 + state_t2 ~ phen_t1 + state_t1

# covariances
phen_t1 ~~ state_t1
phen_t2 ~~ state_t2

# autocorr
phen_t1 ~~ phen_t1
state_t1 ~~ state_t1
phen_t2 ~~ phen_t2
state_t2 ~~ state_t2

# covs
phen_t2 ~ sex_boy + ethn_other + ethn_european + mat_edu_num

state_t2 ~ puberty + agediff_t2 + agemri_t2

state_t1 ~ agediff_t1 + agemri_t1

'

m_mdt1 <- gsub("state", "mdt1", m_general)
m_mdt2 <- gsub("state", "mdt2", m_general)
m_mdt3 <- gsub("state", "mdt3", m_general)
m_mdt4 <- gsub("state", "mdt4", m_general)
m_mdt5 <- gsub("state", "mdt5", m_general)
m_nt <- gsub("state", "nt", m_general)

```


### Specific model specifications 

each state and NT in association with each syndrome scale

```{r}

# create empty lists where to put the model specifications for each measure
# From the general model specifications, we create specific model specifications for each
# syndrome scale. The "var" within the specification will be substituted with each rsfMRI measure

specifications_mdt1 <- list()
specifications_mdt2 <- list()
specifications_mdt3 <- list()
specifications_mdt4 <- list()
specifications_mdt5 <- list()
specifications_nt <- list()



# create vector for the substitution of the var specifications into syndrome scale names 

names <- c("anxdep_", "withdep_", "somatic_", "social_", "thought_", "att_", "agg_", "rulebr_")


# loop to create the various model specifications per syndrome scale 

for (i in names){
  specifications_mdt1[i] <- gsub("phen_", i, m_mdt1) # gsub changes the "var_" string into every element of names in the m_fa specification
  specifications_mdt2[i] <- gsub("phen_", i, m_mdt2)
  specifications_mdt3[i] <- gsub("phen_", i, m_mdt3)
  specifications_mdt4[i] <- gsub("phen_", i, m_mdt4)
  specifications_mdt5[i] <- gsub("phen_", i, m_mdt5)
  specifications_nt[i] <- gsub("phen_", i, m_nt)
}


names(specifications_mdt1) <- paste0(names(specifications_mdt1), "mdt1")
names(specifications_mdt2) <- paste0(names(specifications_mdt2), "mdt2")
names(specifications_mdt3) <- paste0(names(specifications_mdt3), "mdt3")
names(specifications_mdt4) <- paste0(names(specifications_mdt4), "mdt4")
names(specifications_mdt5) <- paste0(names(specifications_mdt5), "mdt5")
names(specifications_nt) <- paste0(names(specifications_nt), "nt")

specifications <- c(specifications_mdt1, specifications_mdt2, specifications_mdt3, specifications_mdt4, specifications_mdt5, specifications_nt) # list of 48 as expected


```

### Model evaluation - fit 

```{r}

# fit all models

fits_all <- sapply(specifications, function(x) sem(x, data = dd2, missing = "fiml", fixed.x = F))


# get all model fit measures that you are interested in 

table_fits <- sapply(fits_all, function(x) fitMeasures(x, c("cfi", "tli", "rmsea", "srmr")))

table_fits <- as.data.frame(t(table_fits)) # for saving, transform into dataframe format after transposing (so fit indices are the cols and models the rows)


table_fits$model <- rownames(table_fits) # create new col with the model name 

table_fits <- table_fits[, c(5, 3, 4, 1, 2)] # change order of cols


table_fits$model <- NULL # for rounding we need to have all vars as numeric

table_fits <- round(table_fits, digits = 3) # round to 3 decimals

# rename 

rownames(table_fits) <- gsub("mdt1", "MDT state 1", 
                            gsub("mdt2", "MDT state 2",
                                 gsub("mdt3", "MDT state 3", 
                                      gsub("mdt4", "MDT state 4",
                                           gsub("mdt5", "MDT state 5",
                                                gsub("- nt", "- NT",
                                                  gsub("anxdep", "Anxious/Depressed",
                                                    gsub("withdep", "Withdrawn/Depressed", 
                                                      gsub("somatic", "Somatic complaints",
                                                           gsub("social", "Social problems",
                                                                gsub("thought", "Thought problems",
                                                                     gsub("att", "Attention problems",
                                                                          gsub("agg", "Aggressive behaviors", 
                                                                               gsub("rulebr", "Rule-breaking behaviors",
                                                     gsub("_", " - ", 
                                                               rownames(table_fits))))))))))))))))






write.csv(table_fits, paste0(tab, "model_fits_syndromes.csv"))

```
all model fits are good as based on absolute model fit indices 


### Output table

Since the fit is good, we can extract the estimates now. 

This part allows us to create a table with all results, that can be used in-text
```{r}

########
# Get estimates 
########

# get standardised estimates 

output_std <- lapply(fits_all, function(x) standardizedSolution(x)) # get std estimates, ci and p vals info 

output_std2 <- as.data.frame(output_std) # change format for saving 

output_std3 <- output_std2[2:3, ] # get the lagged coefficients


#### create output table for in-text ####

output_std3$FDR <- as.numeric("NA")

names(output_std3)

output_table <- data.frame(model = 1:96, 
                           lhs = "NA",
                           op = "NA", 
                           rhs = "NA", 
                           est.std = as.numeric("NA"),
                           se = as.numeric("NA"), 
                           pvalue = as.numeric("NA"), 
                           fdr = as.numeric("NA"))



cols <- paste0("output_std3$", names(output_std3))

cols_p <- grep(".pvalue", cols, value = T)

cols_p <- paste(cols_p, collapse = ", ")

output_table$pvalue <- c(output_std3$anxdep_mdt1.pvalue, output_std3$withdep_mdt1.pvalue, output_std3$somatic_mdt1.pvalue, output_std3$social_mdt1.pvalue, output_std3$thought_mdt1.pvalue, output_std3$att_mdt1.pvalue, output_std3$agg_mdt1.pvalue, output_std3$rulebr_mdt1.pvalue, output_std3$anxdep_mdt2.pvalue, output_std3$withdep_mdt2.pvalue, output_std3$somatic_mdt2.pvalue, output_std3$social_mdt2.pvalue, output_std3$thought_mdt2.pvalue, output_std3$att_mdt2.pvalue, output_std3$agg_mdt2.pvalue, output_std3$rulebr_mdt2.pvalue, output_std3$anxdep_mdt3.pvalue, output_std3$withdep_mdt3.pvalue, output_std3$somatic_mdt3.pvalue, output_std3$social_mdt3.pvalue, output_std3$thought_mdt3.pvalue, output_std3$att_mdt3.pvalue, output_std3$agg_mdt3.pvalue, output_std3$rulebr_mdt3.pvalue, output_std3$anxdep_mdt4.pvalue, output_std3$withdep_mdt4.pvalue, output_std3$somatic_mdt4.pvalue, output_std3$social_mdt4.pvalue, output_std3$thought_mdt4.pvalue, output_std3$att_mdt4.pvalue, output_std3$agg_mdt4.pvalue, output_std3$rulebr_mdt4.pvalue, output_std3$anxdep_mdt5.pvalue, output_std3$withdep_mdt5.pvalue, output_std3$somatic_mdt5.pvalue, output_std3$social_mdt5.pvalue, output_std3$thought_mdt5.pvalue, output_std3$att_mdt5.pvalue, output_std3$agg_mdt5.pvalue, output_std3$rulebr_mdt5.pvalue, output_std3$anxdep_nt.pvalue, output_std3$withdep_nt.pvalue, output_std3$somatic_nt.pvalue, output_std3$social_nt.pvalue, output_std3$thought_nt.pvalue, output_std3$att_nt.pvalue, output_std3$agg_nt.pvalue, output_std3$rulebr_nt.pvalue) # length is 96 as expected 

pvals_adj <- p.adjust(output_table$pvalue, method = "fdr") 

output_table$fdr <- pvals_adj



cols_lhs <- gsub(".pvalue", ".lhs", cols_p)


output_table$lhs <- c(output_std3$anxdep_mdt1.lhs, output_std3$withdep_mdt1.lhs, output_std3$somatic_mdt1.lhs, output_std3$social_mdt1.lhs, output_std3$thought_mdt1.lhs, output_std3$att_mdt1.lhs, output_std3$agg_mdt1.lhs, output_std3$rulebr_mdt1.lhs, output_std3$anxdep_mdt2.lhs, output_std3$withdep_mdt2.lhs, output_std3$somatic_mdt2.lhs, output_std3$social_mdt2.lhs, output_std3$thought_mdt2.lhs, output_std3$att_mdt2.lhs, output_std3$agg_mdt2.lhs, output_std3$rulebr_mdt2.lhs, output_std3$anxdep_mdt3.lhs, output_std3$withdep_mdt3.lhs, output_std3$somatic_mdt3.lhs, output_std3$social_mdt3.lhs, output_std3$thought_mdt3.lhs, output_std3$att_mdt3.lhs, output_std3$agg_mdt3.lhs, output_std3$rulebr_mdt3.lhs, output_std3$anxdep_mdt4.lhs, output_std3$withdep_mdt4.lhs, output_std3$somatic_mdt4.lhs, output_std3$social_mdt4.lhs, output_std3$thought_mdt4.lhs, output_std3$att_mdt4.lhs, output_std3$agg_mdt4.lhs, output_std3$rulebr_mdt4.lhs, output_std3$anxdep_mdt5.lhs, output_std3$withdep_mdt5.lhs, output_std3$somatic_mdt5.lhs, output_std3$social_mdt5.lhs, output_std3$thought_mdt5.lhs, output_std3$att_mdt5.lhs, output_std3$agg_mdt5.lhs, output_std3$rulebr_mdt5.lhs, output_std3$anxdep_nt.lhs, output_std3$withdep_nt.lhs, output_std3$somatic_nt.lhs, output_std3$social_nt.lhs, output_std3$thought_nt.lhs, output_std3$att_nt.lhs, output_std3$agg_nt.lhs, output_std3$rulebr_nt.lhs)



cols_rhs <- gsub(".lhs", ".rhs", cols_lhs)

output_table$rhs <- c(output_std3$anxdep_mdt1.rhs, output_std3$withdep_mdt1.rhs, output_std3$somatic_mdt1.rhs, output_std3$social_mdt1.rhs, output_std3$thought_mdt1.rhs, output_std3$att_mdt1.rhs, output_std3$agg_mdt1.rhs, output_std3$rulebr_mdt1.rhs, output_std3$anxdep_mdt2.rhs, output_std3$withdep_mdt2.rhs, output_std3$somatic_mdt2.rhs, output_std3$social_mdt2.rhs, output_std3$thought_mdt2.rhs, output_std3$att_mdt2.rhs, output_std3$agg_mdt2.rhs, output_std3$rulebr_mdt2.rhs, output_std3$anxdep_mdt3.rhs, output_std3$withdep_mdt3.rhs, output_std3$somatic_mdt3.rhs, output_std3$social_mdt3.rhs, output_std3$thought_mdt3.rhs, output_std3$att_mdt3.rhs, output_std3$agg_mdt3.rhs, output_std3$rulebr_mdt3.rhs, output_std3$anxdep_mdt4.rhs, output_std3$withdep_mdt4.rhs, output_std3$somatic_mdt4.rhs, output_std3$social_mdt4.rhs, output_std3$thought_mdt4.rhs, output_std3$att_mdt4.rhs, output_std3$agg_mdt4.rhs, output_std3$rulebr_mdt4.rhs, output_std3$anxdep_mdt5.rhs, output_std3$withdep_mdt5.rhs, output_std3$somatic_mdt5.rhs, output_std3$social_mdt5.rhs, output_std3$thought_mdt5.rhs, output_std3$att_mdt5.rhs, output_std3$agg_mdt5.rhs, output_std3$rulebr_mdt5.rhs, output_std3$anxdep_nt.rhs, output_std3$withdep_nt.rhs, output_std3$somatic_nt.rhs, output_std3$social_nt.rhs, output_std3$thought_nt.rhs, output_std3$att_nt.rhs, output_std3$agg_nt.rhs, output_std3$rulebr_nt.rhs)

output_table$op <- "~"


cols_est.std <- gsub(".rhs", ".est.std", cols_rhs)


output_table$est.std <- c(output_std3$anxdep_mdt1.est.std, output_std3$withdep_mdt1.est.std, output_std3$somatic_mdt1.est.std, output_std3$social_mdt1.est.std, output_std3$thought_mdt1.est.std, output_std3$att_mdt1.est.std, output_std3$agg_mdt1.est.std, output_std3$rulebr_mdt1.est.std, output_std3$anxdep_mdt2.est.std, output_std3$withdep_mdt2.est.std, output_std3$somatic_mdt2.est.std, output_std3$social_mdt2.est.std, output_std3$thought_mdt2.est.std, output_std3$att_mdt2.est.std, output_std3$agg_mdt2.est.std, output_std3$rulebr_mdt2.est.std, output_std3$anxdep_mdt3.est.std, output_std3$withdep_mdt3.est.std, output_std3$somatic_mdt3.est.std, output_std3$social_mdt3.est.std, output_std3$thought_mdt3.est.std, output_std3$att_mdt3.est.std, output_std3$agg_mdt3.est.std, output_std3$rulebr_mdt3.est.std, output_std3$anxdep_mdt4.est.std, output_std3$withdep_mdt4.est.std, output_std3$somatic_mdt4.est.std, output_std3$social_mdt4.est.std, output_std3$thought_mdt4.est.std, output_std3$att_mdt4.est.std, output_std3$agg_mdt4.est.std, output_std3$rulebr_mdt4.est.std, output_std3$anxdep_mdt5.est.std, output_std3$withdep_mdt5.est.std, output_std3$somatic_mdt5.est.std, output_std3$social_mdt5.est.std, output_std3$thought_mdt5.est.std, output_std3$att_mdt5.est.std, output_std3$agg_mdt5.est.std, output_std3$rulebr_mdt5.est.std, output_std3$anxdep_nt.est.std, output_std3$withdep_nt.est.std, output_std3$somatic_nt.est.std, output_std3$social_nt.est.std, output_std3$thought_nt.est.std, output_std3$att_nt.est.std, output_std3$agg_nt.est.std, output_std3$rulebr_nt.est.std)


cols_se <- gsub(".est.std", ".se", cols_est.std)


output_table$se <-  c(output_std3$anxdep_mdt1.se, output_std3$withdep_mdt1.se, output_std3$somatic_mdt1.se, output_std3$social_mdt1.se, output_std3$thought_mdt1.se, output_std3$att_mdt1.se, output_std3$agg_mdt1.se, output_std3$rulebr_mdt1.se, output_std3$anxdep_mdt2.se, output_std3$withdep_mdt2.se, output_std3$somatic_mdt2.se, output_std3$social_mdt2.se, output_std3$thought_mdt2.se, output_std3$att_mdt2.se, output_std3$agg_mdt2.se, output_std3$rulebr_mdt2.se, output_std3$anxdep_mdt3.se, output_std3$withdep_mdt3.se, output_std3$somatic_mdt3.se, output_std3$social_mdt3.se, output_std3$thought_mdt3.se, output_std3$att_mdt3.se, output_std3$agg_mdt3.se, output_std3$rulebr_mdt3.se, output_std3$anxdep_mdt4.se, output_std3$withdep_mdt4.se, output_std3$somatic_mdt4.se, output_std3$social_mdt4.se, output_std3$thought_mdt4.se, output_std3$att_mdt4.se, output_std3$agg_mdt4.se, output_std3$rulebr_mdt4.se, output_std3$anxdep_mdt5.se, output_std3$withdep_mdt5.se, output_std3$somatic_mdt5.se, output_std3$social_mdt5.se, output_std3$thought_mdt5.se, output_std3$att_mdt5.se, output_std3$agg_mdt5.se, output_std3$rulebr_mdt5.se, output_std3$anxdep_nt.se, output_std3$withdep_nt.se, output_std3$somatic_nt.se, output_std3$social_nt.se, output_std3$thought_nt.se, output_std3$att_nt.se, output_std3$agg_nt.se, output_std3$rulebr_nt.se)


output_table$model <- paste0(output_table$lhs, " ", output_table$op, " ", output_table$rhs)

out <- output_table[ , c(1, 5, 6, 7, 8)] 



out2 <- out
rownames(out2) <- out2$model 
out2$model <- NULL

out2 <- round(out2, digits = 3)

# separate into temporal direction (brain --> beh; beh --> brain)

BrtoBeh <- grep("mdt1_t1|mdt2_t1|mdt3_t1|mdt4_t1|mdt5_t1| nt_t1", rownames(out2), value = T)

out_BrtoBeh <- out2[rownames(out2) %in% BrtoBeh, ]
out_BehtoBr <- out2[rownames(out2) %notin% BrtoBeh, ]


out2 <- rbind(out_BrtoBeh, out_BehtoBr)


rownames(out2) <- gsub("mdt1", "MDT state 1", 
                            gsub("mdt2", "MDT state 2",
                                 gsub("mdt3", "MDT state 3", 
                                      gsub("mdt4", "MDT state 4",
                                           gsub("mdt5", "MDT state 5",
                                                gsub("nt ", "NT ",           gsub("anxdep", "Anxious/Depressed",
                                                 gsub("withdep", "Withdrawn/Depressed", 
                                                      gsub("somatic", "Somatic complaints",
                                                           gsub("social", "Social problems",
                                                                gsub("thought", "Thought problems",
                                                                     gsub("att", "Attention problems",
                                                                          gsub("agg", "Aggressive behaviors", 
                                                                               gsub("rulebr", "Rule-breaking behaviors",
                                                                                    gsub("_t2", " T2", 
                                                                                         gsub("_t1", " T1",
                                                                                         rownames(out2)))))))))))))))))








write.csv(out2, paste0(tab, "results_syndromes.csv"))


```


